---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sigminerUtils

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/sigminerUtils)](https://CRAN.R-project.org/package=sigminerUtils)
[![Codecov test coverage](https://codecov.io/gh/selkamand/sigminerUtils/branch/master/graph/badge.svg)](https://app.codecov.io/gh/selkamand/sigminerUtils?branch=master)
[![R-CMD-check](https://github.com/selkamand/sigminerUtils/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/selkamand/sigminerUtils/actions/workflows/R-CMD-check.yaml)
![GitHub Issues or Pull Requests](https://img.shields.io/github/issues-closed/selkamand/sigminerUtils)
[![code size](https://img.shields.io/github/languages/code-size/selkamand/sigminerUtils.svg)](https://github.com/selkamand/sigminerUtils)
![GitHub last commit](https://img.shields.io/github/last-commit/selkamand/sigminerUtils)
<!-- badges: end -->


> [!WARNING]  
> This package is in early development and not ready for use


Utility package to simplify the process of running sigminer, storing signature results and producing key visualisations

## Installation

You can install the development version of sigminerUtils like so:

``` r
# install.packages('remotes')
remotes::install_github('selkamand/sigminerUtils')
```

## Usage

Load required packages

```{r}
library(sigminerUtils)
```

### Run in Single Sample Mode from Oncoanalyser outputs

sigminerUtils was primarily designed to support mutation files (SNV, SV, Copynumber files) output by [oncoanalyser](https://github.com/nf-core/oncoanalyser).

```{r include=FALSE}

# Ensure folders we are going to create don't already exist
unlink("pcawg_signature_results", recursive = TRUE)
unlink("colo829_signature_results", recursive = TRUE)
unlink("pcawg_reference_set", recursive = TRUE)
unlink("colo829_signature_results_with_refset", recursive = TRUE)

```

```{r echo=TRUE, results='hide', message=FALSE, warning=FALSE}
path_snvs <- system.file("colo829_testfiles/COLO829v003T.purple.somatic.vcf.gz", package = "sigminerUtils")
path_cnvs <- system.file("colo829_testfiles/COLO829v003T.purple.cnv.somatic.tsv", package = "sigminerUtils")
path_svs <-  system.file("colo829_testfiles/COLO829v003T.purple.sv.vcf.gz", package = "sigminerUtils")

sig_analyse_mutations_single_sample_from_files(
  sample_id = "COLO829v003T", 
  vcf_snv = path_snvs, 
  segment = path_cnvs, 
  vcf_sv = path_svs,
  include = "pass",
  ref = "hg38",
  cores = 4,
  output_dir = "colo829_signature_results"
)
```


### Run in Cohort Mode using a manifest file

Run on a cohort of samples using a [manifest file](inst/example_manifest.tsv).

> [!NOTE]  
> See `?sig_analyse_cohort_from_files()` for details on the expected manifest format

```{r}
# Extract PCAWG data into working directory
pcawg_tar = system.file("pcawg/pcawg_data.tar.gz", package = "sigminerUtils")
untar(tarfile = pcawg_tar, exdir = "pcawg")

# Supply a manifest file
path_manifest <- system.file("pcawg/manifest.tsv", package = "sigminerUtils")

# Run the signature analysis
sig_analyse_cohort_from_files(
  manifest = path_manifest,
  ref = "hg19",
  output_dir = "pcawg_signature_results", 
  cores=1, 
  verbose=FALSE
) 
```

### Create a reference set

Single sample mutational signature analyses are much more informative when you
have a reference set of samples to compare against.

A reference set allows a single sample analysis to include

1. Similar sample analysis
2. UMAPs

```{r}
sig_create_reference_set(
  "pcawg_signature_results", 
  outfolder = "pcawg_reference_set", 
  create_umaps = TRUE, 
  umap_n_neighbours = 2,
  verbose=FALSE
)
```


### Rerun Analysis using Reference Sets

Once a reference set has been produced, it can be used in any single sample / cohort analysis to produce additional outputs (umap_plot data + similar sample analysis)

```{r echo = TRUE, results='hide', message=FALSE, warning=FALSE}
path_snvs <- system.file("colo829_testfiles/COLO829v003T.purple.somatic.vcf.gz", package = "sigminerUtils")
path_cnvs <- system.file("colo829_testfiles/COLO829v003T.purple.cnv.somatic.tsv", package = "sigminerUtils")
path_svs <-  system.file("colo829_testfiles/COLO829v003T.purple.sv.vcf.gz", package = "sigminerUtils")


sig_analyse_mutations_single_sample_from_files(
  sample_id = "COLO829v003T", 
  vcf_snv = path_snvs, 
  segment = path_cnvs,
  vcf_sv = path_svs,
  ref_tallies = "pcawg_reference_set/refmatrix.tally.parquet",
  ref_umaps_prefix = "pcawg_reference_set/refmatrix.umaps",
  include = "pass",
  ref = "hg38",
  output_dir = "colo829_signature_results_with_refset", 
  cores = 4, # Run with 4 cores
  verbose = FALSE
)
```

### Parse Results into SigStory object

```{r}
sigstory = sigminer2sigstory(
  signature_folder = "colo829_signature_results_with_refset/COLO829v003T/"
)
```

