---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sigminerUtils

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/sigminerUtils)](https://CRAN.R-project.org/package=sigminerUtils)
<!-- badges: end -->


Utility package to simplify the process of running sigminer, storing signature results and producing key visualisations

*Warning:* This package is in early development and not ready for use


## Installation

You can install the development version of sigminerUtils like so:

``` r
# install.packages('remotes')
remotes::install_github('selkamand/sigminerUtils')
```

## Usage

1. Start by creating a Signature database

```{r, eval = FALSE}
sig_create_database('./my_sig_database.sqlite', overwrite = TRUE)
```


2. Run Signature Analysis

```{r, eval = FALSE}
# Variant Files
path_maf <-  system.file(package = "sigminerUtils", path = "test.maf") 
path_copynumber <- system.file(package = "sigminerUtils", path = "test.copynumber.csv") 
path_sv <- system.file(package = "sigminerUtils", path = "test.sv.csv") 

sig_analyse_mutations(
  maf = path_maf, 
  copynumber = read.csv(path_copynumber),
  structuralvariant = read.csv(path_sv),
  ref = "hg19", 
  output_dir = "./signatures"
  )
```

3. Create a reference matrix

```{r}
sig_create_reference_set(
  path_to_signature_directory = "./signatures",
  outfolder = "./reference"
  )
```


``
4. Use reference matrix to rerun your signature analysis
```{r}
# Variant Files
path_maf <-  system.file(package = "sigminerUtils", path = "test.maf") 
path_copynumber <- system.file(package = "sigminerUtils", path = "test.copynumber.csv") 
path_sv <- system.file(package = "sigminerUtils", path = "test.sv.csv") 

# Database of reference of sample catalogues
path_reference_tallies <- system.file(package = "sigminerUtils", path = "reference/refmatrix.tally.parquet")

sig_analyse_mutations(
  maf = path_maf, 
  copynumber = read.csv(path_copynumber),
  structuralvariant = read.csv(path_sv),
  ref = "hg19", 
  ref_tallies = path_reference_tallies
  output_dir = "./signatures"
  )

```


3. Load results from output folder into the database

```{r, eval = FALSE}
sig_add_to_database('./signatures', './my_sig_database.sqlite', ref = "hg19")
```
Note you can optionally supply sample metadata

```{r, eval = FALSE}
sample_metadata <- data.frame(
  sampleId = c('TCGA-A2-A0T5-01', 'TCGA-CA-6717-01', 'TCGA-CF-A9FF-01'),
  disease = c('Breast', 'Colorectal', 'Bladder'),
  description = character(3)
)

sig_add_to_database('./signatures', './my_sig_database.sqlite', ref = "hg19", metadata = sample_metadata)
```


